---
title: "PS03: Hedonic House Price Regression"
author: "ANSWER KEY"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## README

Complete the following questions below and show all work.
You must submit a compiled R report with all code and answers in this document.
Compile the report to html or PDF and submit on Canvas.
html is preferred.
Make sure to write your name above where it says “WRITE YOUR NAME HERE”.
Include it within the quotation marks.

Before writing answering the following questions,
make sure to compile the report to make sure that everything is working properly.
As you answer each question,
compile after each question to make sure you are not making any errors.
If you are experiencing an error, read the error carefully.
One tactic is to selectively delete questions/sections
until the report compiles.
This will help you narrow down where the coding error is occurring.

Run the following code chunk to set up your R environment with the 
necessary packages:
```{r}
#| message: false
# Load tidyverse packages
library(tidyverse)

# Load broom for tidy() regression outputs
if (!require("broom")) install.packages("broom") 
library(broom) 

# Load AER package (install if not available)
if (!require("AER")) install.packages("AER")
```

## DATA

Run `data("HousePrices")` to load in the data object `HousePrices` 
from the `AER` package.
```{r}
data("HousePrices")
```

These data originally come from the paper
[*Semiparametric Estimation of a Hedonic Price Function*](https://www.jstor.org/stable/2285156) by Paul Anglin and Ramazan Gencay (1996) 
and were digitized for the econometrics textbook by [Verbeek](https://thenigerianprofessionalaccountant.files.wordpress.com/2013/04/modern-econometrics.pdf) (2004).

Familiarize yourself with the data set by using `?HousePrices` to view the 
documentation page with variable definitions.
You should also take a look through the data by using functions like 
`head()`, `glimpse()`, or `View()`.

## Academic Integrity

You are allowed to consult the lectures, notes, problem sets, quizzes, and the internet on this assignment. 
However, this is an individual assessment of your own understanding 
so you are are not permitted to collaborate with anyone else.
You may consult Stackoverflow, ChatGPT, other LLM models, etc.
for help with R coding, but you are responsible for understanding any code 
which you use to answer these questions.
All written answers must be your in your own words. 

*If you are suspected of plagiarizing, you may recieve a zero for this assignment and possibly for the course.*

## Question 1

Before you run any regressions, I want you to look for any visual patterns in the data.
We will first start with a scatterplot of the prices observed in this sample to get an idea of the shape of the distribution.
I have provided the basic code to get a default `ggplot` histogram below,
but please feel free to add to the code to adjust the labels, aesthetics, 
binwidth, add a title, etc. to your preferences.

```{r price_histogram}
HousePrices %>% 
  ggplot(aes(x=price)) +
  geom_histogram(
    fill = "#2D4F67", # Set a custom color by hex code
    binwidth = 7000   # set a custom width of the bars
    ) +
  labs( # set custom labels for title, x axis, and y axis
    title = "Windsor Canada Housing Price Distribution",
    subtitle = "Data from Anglin and Gencay (1996)",
    x = "Sale Price",
    y = "Number of Houses"
  ) +
  theme( plot.title = element_text(hjust = 0.5) ) # center title text
```

### **1a:** What patterns do you see in the distribution of house prices? *(10 pts)*

**Now take a look through the other variables in the dataset using `?AER::HousePrices`
or `colnames(HousePrices)`. 
Choose one variable which you think is (intuitively) an important factor in price.**

### **1b:** Describe the variable you chose. What are the units? Is it continuous or discrete? *(6 pts)*

- **If you chose a *continuous* variable, then fill in the blank in the r chunk labeled `continuous_scatter` below with your variable to see a scatterplot.**

```{r continuous_scatter}
HousePrices %>%
  ggplot(aes( x = lotsize, y = price )) +
  geom_point(color = "#2D4F67") + # set custom dot color
  labs( # set custom labels for title, x axis, and y axis
    title = "Houses' Lot Size vs. Sale Price",
    x = "Lot Size (square feet)",
    y = "Sale Price"
  ) +
  theme( plot.title = element_text(hjust = 0.5) ) # center title text
```

- **If you chose a *discrete* variable (either an indicator or categorical one),
then it might be better to use a 'ridgeline' plot where the price density at 
each level of your explanatory variable is plotted as its own 'ridge'.
Fill in the blank in the code chunk `discrete_ridgeline` with your variable inside of `as.factor()` so that R knows to interpret it as a factor, not numeric.**

```{r discrete_ridgeline}
if (!require("ggridges")) install.packages("ggridges")
HousePrices %>%
  ggplot(aes( x = price, y = as.factor(bedrooms) )) +
  ggridges::geom_density_ridges(aes( fill = bedrooms )) +
  labs( # set custom labels for title, x axis, and y axis
    title = "Houses' Lot Size vs. Number of Bedrooms",
    x = "Sale Price",
    y = "Bedrooms",
  ) +
  theme( plot.title = element_text(hjust = 0.5) ) + # center title text
  guides( fill = "none" )
```

### **1c:** Describe some patterns you see in the plot above between price and your chosen variable. *(10 pts)*

## Question 2

### **2a:** Write out a regression model mathematically with price as the dependent variable and the same independent variable you used above: *(8 pts)*

$$
\text{price}_i = \beta_0 + \beta_1 \text{bedrooms}_i + u_i
$$

### **2b:** Now write out a hypothesis test that the regression above will be used to test *(6 pts)*

$$
H_0: \beta_1 = 0 \\
H_1: \beta_1 \neq 0
$$

### **2c:** Run your hypothesis test at the 95% significance level. Interpret your results *(12 pts)*
**Include all code you use in its own R chunk**
```{r simple_regression}
model1 <- lm(data = HousePrices, price ~ bedrooms) %>% 
  tidy()
model1
```

For every additional bedroom in a house, 
the sale price is expected to increase by \$`r model1$estimate[2]`.
The $t$-statistic for the two-tailed hypothesis test in question 2b
is `r model1$statistic[2]` which is less than the 

## Question 3

You will repeat the steps in Question 2, but this time using a 
*multiple linear regression model* with *two explanatory variables*.
Pick another variable other than `price` or the variable you chose in Question 1.

### **3a:** Write out your regression model from 2a but with an additional explanatory variable *(8 pts)*
$$
\text{price}_i = \beta_0 + \beta_1\text{variable1}_i + \beta_2\text{variable2}_i + u_i
$$

### **3b:** Run this regression and report your results *(12 pts)*

```{r multiple_regression}
# your code goes here
summary(lm(price ~ bedrooms, data = HousePrices))$r.squared
summary(lm(price ~ bedrooms, data = HousePrices))$adj.r.squared
```

**Make sure to address the following prompts:**

- How does your estimate of $\beta_1$ change between `simple_regression` and `multiple_regression`? What about the standard error?

- How does the $R^2$ value change? What about the *adjusted* $R^2$? 
Is this what you would predict from what you know about those measures?

## Question 4

You will repeat the steps in Question 3, but this time using the
*multiple linear regression model* with *all* of the variables besides price included. 
We'll call this the *full model*:

$$
\begin{align}
\text{price}_i = & \beta_0 + \beta_1 \text{lotsize}_i + \beta_2 \text{bedrooms}_i + \\
 & \beta_3 \text{bathrooms}_i + \beta_4 \text{stories}_i + \beta_5 \text{driveway}_i + \\
 & \beta_6 \text{recreation}_i + \beta_7 \text{fullbase}_i + \beta_8 \text{gasheat}_i + \\
 & \beta_9 \text{aircon}_i + \beta_{10} \text{garage}_i + \beta_{11} \text{prefer}_i + u_i
\end{align}
$$

### **4a:** Run the full model and interpret the $\beta$ on the same variable you chose in 1b *(18 pts)*

```{r full_regression}
# your code goes here
```

- How does your estimate of that $\beta$ change between `multiple_regressio` and `full_regression`? What about the standard error?

- How does the $R^2$ value change? What about the *adjusted* $R^2$? 

## Compile and submit your work *(10 pts)*

Make sure that you have written answers to all 7 questions 
and that all code you used in your answers in contained in code chunks.
Then compile your Rmd file using the 'Knit' button at the top of the source pane
or with the hotkey `Shift Cmd/Ctrl K`.
Open the `PS03.html` file that will appear in the same directory as your 
`PS03.Rmd` file and check to make sure that the output looks correct.

Once you are sure that your work is correct and formatted correctly,
upload this file to the assignment named 'ps03' on Canvas.